Sketch Recognition using Enhanced Particle Swarm Optimization

### 2010-12-10 trunk/Swarm/swarmModified.tex

%%\usepackage{icgst}
\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}
%\usepackage{named}
%\usepackage{hyperref}
%\usepackage[dvips,pdftex]{hyperref}
%\usepackage{subfigure}
\usepackage{array}
\usepackage{amssymb}
%\usepackage{epsfig}
%\usepackage{epstopdf}
%\usepackage{xltxtra}
\usepackage{graphicx}
\usepackage{sectsty}
\sectionfont{\normalsize  }
\subsectionfont{ \normalsize\it \textbf }
\subsubsectionfont{\normalsize \it \textbf }
%\paragraphfont{\large\it}
%\subparagraphfont{\normalsize\it}

@@

% \biboptions{}
\setlength\hoffset{0in}
% \setlength\topmargin{0in}
% \setlength\headheight{0in}
% \setlength\headsep{-0.5in}
% \setlength\textheight{10.0in}
 \setlength\textwidth{7in}
 \setlength\oddsidemargin{-0.5in}
 \setlength\evensidemargin{-0.5in}
% \setlength\parindent{0.25in}
% \setlength\parskip{0.25in} 
\linespread{1.5}

\begin{document}

@@

% \twocolumn[
% \scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
% \title{Sketch Recognition using Enhanced Particle Swarm Optimization}
% \vspace{0.1in}
% \author{}  
% \begin{center}
% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek abouldahab\\
% \textit{Electronic Research Institute,} Cairo, Egypt,\\
% \textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
% {melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
% %\texttt{http://www.faculty.university.edu.any}
% \end{center}
% \vspace{0.25in} 
% ]

% \twoauthors
% 
%  {Maha El Meseery, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg} 
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

@@

\author{
\small{
Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
 \textsuperscript{a} \small{Signals Processing Group, Computers and Systems Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
  \small {\textsuperscript{b}Cairo Metro Company ,
   Ministry of transport , Cairo, Egypt.}\\
   \small{melemseery@eri.sci.eg,  mafakhr@mcit.gov.eg, heshoaboda@hotmail.com
   }
}
\date{}
% {\sthanks{Corresponding Author},}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
%   {Tarek abouldahab} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
%  {Maha El Meseery \sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

\maketitle
 \begin{abstract}

%This study 
Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that the enhanced particle swarm optimization improves the final recognition accuracy.

\end{abstract}

@@

This paper modifies the sketch recognition system presented in \cite{mypaper}, to use enhanced particle swarm algorithm (EPSO) to optimally segment user's strokes. As in the Original system the users can draw symbols using any number of strokes in any order they like. The system segments each stroke using two different EPSO algorithms. Segmentation is based on dividing the digital curve into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our system uses a second segmentation EPSO algorithm to segment strokes into a set of lines and circular arcs. After segmentation, a symbol features vector composed of different statistical, geometrical and spatial features is deduced. The final classification step, SVM classifier uses the computed feature vector to classify the segments into one of the previously trained classes.

@@

\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. Figure \ref{fig:flowChart} shows the flow chart of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.  The following section provides a detailed description of each step. The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
 \begin{figure}
	\centering
		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
	\caption{System flow chart} The chart shows the flow of process and data processed thought the system. 
	\label{fig:flowChart}
\end{figure}

\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

After the system computes all the speed, time difference and curvature information it proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions $Region_i$; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process for curvature, time difference and speed information. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points. % (as shown are redundant)
\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{ellipsefit}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two further segmentation algorithms that divide the stroke to either lines or lines and curves. After two different segmentations are generated, the system chooses the segmentation with the minimum error. The following sections describe in details the ellipse detection algorithm and the two segmentation algorithms used to divide the stroke. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}

@@

	 \includegraphics[scale=0.7]{pso1.jpg}			
	\caption{DPSO algorithm encoding}
	\label{fig:pso1}
\end{figure}
Two DPSO algorithms are used to generate different segmentations for each stroke the user draws. For each segmentation generated, an error is evaluated. The segmentation with the minimum error value is chosen as the best stroke segmentation (Figure \ref{fig:segblock}). The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. 
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem using a binary array with the same length $N$ as the input stroke.

@@

	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
	\end{description}
\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
%\\
The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$. The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed using equation %\ref{eq:fitnessSwarm1}
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution. Otherwise the system favors the lower number of vertices.

@@

\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

\cite{mypaper}  
\section{Experiments} 
\label{sec:Experiments}
In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. Five different splits were generated to divide the dataset into training and test sets. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.

>>>>

%%\usepackage{icgst}
%\documentclass[12pt,times,a4paper]{article}%[10pt]{article}
%% Use the option review to obtain double line spacing
 \documentclass[preprint,1p,times,review]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}
%\usepackage{named}
%\usepackage{hyperref}
%\usepackage[dvips,pdftex]{hyperref}
%\usepackage{subfigure}
\usepackage{array}
\usepackage{amssymb}
%\usepackage{epsfig}
%\usepackage{epstopdf}
%\usepackage{xltxtra}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{sectsty}
%\sectionfont{\normalsize  }
%\subsectionfont{ \normalsize\it \textbf }
%\subsubsectionfont{\normalsize \it \textbf }
%\paragraphfont{\large\it}
%\subparagraphfont{\normalsize\it}

@@

% \biboptions{}
% \setlength\hoffset{0in}
% % \setlength\topmargin{0in}
% % \setlength\headheight{0in}
% % \setlength\headsep{-0.5in}
% % \setlength\textheight{10.0in}
%  \setlength\textwidth{7in}
%  \setlength\oddsidemargin{-0.5in}
%  \setlength\evensidemargin{-0.5in}
% % \setlength\parindent{0.25in}
% % \setlength\parskip{0.25in}
% \linespread{1.5}
\begin{abstract}

%This study
%%%This study

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Algorithms to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. A comparision is presented between using Traditional Particle Swarm Optimization and the Enhanced Particle Swarm Optimization (EPSO) which is based on the perturbation of the particle velocity. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS). Results also show that using an Enhancment on the Tradiontal Swarm algorithm improves recognition rate.

\end{abstract}
\begin{document}

@@

% \twocolumn[
% \scalebox{1.00}{\includegraphics[width=1.00\textwidth]{aimlhead}}
% \title{Sketch Recognition using Enhanced Particle Swarm Optimization}
% \vspace{0.1in}
% \author{}
% \begin{center}
% Maha El Meseery\footnote{Corresponding Author }, Mahmoud Fakhr El Din, Tarek
%abouldahab\\
% \textit{Electronic Research Institute,} Cairo, Egypt,\\
% \textit{ Cairo Metro Company, Ministry of transport }Cairo, Egypt\\
% {melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg,  heshoaboda@hotmail.com},\\
% %\texttt{http://www.faculty.university.edu.any}
% \end{center}
% \vspace{0.25in}
% ]

% \twoauthors
%
%  {Maha El Meseery, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
%\\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

@@

\author{
\small{
Maha El Meseery \footnote{Corresponding Author} \textsuperscript{a}, Mahmoud
Fakhr El Din \textsuperscript{a}, Tarek abouldahab \textsuperscript{b}}\\
 \textsuperscript{a} \small{Signals Processing Group, Computers and Systems
Department,}\\ \small{ Electronic Research Institute, Cairo, Egypt.}\\
  \small {\textsuperscript{b}Cairo Metro Company ,
   Ministry of transport , Cairo, Egypt.}\\
   \small{melemseery@eri.sci.eg,  mafakhr@mcit.gov.eg, heshoaboda@hotmail.com
   }
}
\date{}
% {\sthanks{Corresponding Author},}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
%   {Tarek abouldahab} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}
%  {Maha El Meseery \sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
% 	 {Tarek abouldahab\sthanks{Corresponding Author}} {Cairo Metro Company
%\\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

\maketitle
%  \begin{abstract}
%
% %This study
% %%%This study
%
% Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is developing a sketch recognition system that facilitates using sketches in computer systems. The system uses Particle Swarm Optimization (PSO) algorithm to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Two different segmentation algorithms were evaluated and results show that both algorithms improve the final recognition results. Experiments were conducted on three different dataset, simple presentation (Hs-DS), Electrical Design (EL-DS) and Logic Design (LD-DS) symbols. Results show that the effectiveness of the proposed system on various dataset especially on datasets with low shape diversity as in Logic Design dataset (LD-DS).
%
%
%  \textsl{Sketches and drawings are widely used as a simple method of expressing
% thought and designs. The goal of this paper is to improve our sketch recognition
% system by using an enhanced particle swarm optimization algorithm. The Enhanced
% Particle Swarm Optimization (EPSO) is based on the perturbation of the particle
% velocity is used to optimally segment the strokes the user draws into meaningful
% geometric primitives. These geometric primitives are grouped to formulate
% symbols which are further identified using a Support Vector Machines (SVM)
% classifier. Experiments were conducted on the benchmark dataset of simple
% presentation (Hs-DS) symbols. Results show that the enhanced particle swarm
% optimization improves the final recognition accuracy.}
% \end{abstract}

@@

In PR based methods, systems compute properties that describe symbols then use a
classification method to recognize symbols. The properties used can either be
based on gesture properties, global shape properties or image based properties.
Gesture properties are set of features that represent how the user draws a
symbol not how it looks like (e.g., identify a rectangle when a user draws a L
corner). In gesture properties approach, systems identify symbols by using
gesture properties of some of its underlying strokes. A system that uses
gestures properties assumes that users draws symbols in specific order and style
\cite{gestureexample12,aideddesgin22}. In global shape properties approach,
systems investigate general shape properties and its underlying strokes. Using
shape properties (e.g., ratio of the area of convex hull to the area of shape)
relaxes the assumption and the role of each stroke in the symbol but it does not
truly represent the appearance or fine details of the symbol
\cite{DiagramOfflineConvexHull,Cali63}. Unlike shape properties based methods,
image based methods disregard individual strokes and focus on the overall
appearance of shapes. These system convert the shapes into image and extracted
image based features (e.g. local gradient
feature)\cite{Oltmans07,imagetrainable48}. In \cite{imagetrainable48}, they used
four different distance measures to match user symbol with the prototype
database. The prototype of a shape is represented as an image of the symbol that
is normalized and scaled.  However, the problem is that the shape prototype must
represent all transformation and variation the input shape can have.   %Image
based methods are used in \cite{Oltmans07,imagetrainable48} where they computed
a code book of 1000 features to recognize symbols.

Methods based on gesture and global shape properties produce systems that are
dependent on assumptions about how strokes can be grouped. These systems assume
that users will finish drawing a shape before moving to the next one
\cite{Cali63,geometrydomain49}. This assumption allows the system to temporally
group strokes and then recognizes them based on their strokes or global shape
properties. Alvarado \cite{AlvaradoDigital} argued that this assumption does not
hold in natural sketches.

In this paper, we choose to implement a hybrid method that includes shape
property, gesture and hierarchical information. Our method attempts to avoid
many of the problems faced by approaches based on individual gestures or global
shape properties while trying to gain the additional information that the stroke
based system uses. Firstly, the system uses particle swarm optimization (PSO) to
optimally segment user's strokes. Users can draw symbols using any number of
strokes in any order they like. The system segments each stroke using two
different PSO algorithms. The first segmentation algorithm is based on dividing
the digital curve into polygons\cite{PolygonApproximationPSO}. Our algorithm
enhances its performance by adding curvature and other local information during
segmenting strokes to achieve better results. To handle complex strokes, our
system uses a second segmentation PSO algorithm that segment strokes into a set
of lines and circular arcs. The two segmentation algorithms can either by used
individually or in parallel then chosing the segmentation that gives the minimum
error. After segmentation, a feature vector composed of different statistical,
geometrical and composite features is deduced. The final classification step
uses a SVM classifier to classify symbols into one of the previously trained
classes based on the features vector computed\cite{mypaper}.

 In addition to the system mentioned above, an Enhnaced Swarm algorithm is compared to the Traditional Particle swarm algorithm\cite{stockPaper}. The Enhanced swarm algorithm is based on chaning the particle velocity by replacing it with the weighted differance between the particle and another random in the space. This random relacing is only used when the new velociy and location is better then the current particle values. The procedure introude new expolration factor the swarm system which helps in reaching the solution faster and more obtimially.

% %This paper modifies the sketch recognition system presented in \cite{mypaper},
% to use enhanced particle swarm algorithm (EPSO) to optimally segment user's
% strokes. As in the Original system the users can draw symbols using any number
% of strokes in any order they like. The system segments each stroke using two
% different EPSO algorithms. Segmentation is based on dividing the digital curve
% into polygon\cite{PolygonApproximationPSO}. To handle complex strokes, our
% system uses a second segmentation EPSO algorithm to segment strokes into a set
% of lines and circular arcs. After segmentation, a symbol features vector
% composed of different statistical, geometrical and spatial features is deduced.
% The final classification step, SVM classifier uses the computed feature vector
% to classify the segments into one of the previously trained classes.

@@

\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
%The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}.

The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition.
Figure \ref{fig:flowChart} shows the block diagram of the system; the first
process is collecting stroke points then computing the curvature and possible
dominant points.  The following section provides a detailed description of each
step. The main block of the system remains the same but there are some
modifications in the segmentation algorithm to improve its efficiency. This study compares between using Traditional PSO and an Enhanced Particle swarm algorithm. Other modifications are explained in the details in
the next sections.
%  \begin{figure}
% 	\centering
% 		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
% 	\caption{System flow chart} The chart shows the flow of process and data processed thought the system.
% 	\label{fig:flowChart}
% \end{figure}

 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/Blockdiagram.png}
	\caption{System Block Diagram} The chart shows the system block diagram.
	\label{fig:flowChart}
\end{figure}

\subsection{Preprocessing}
\label{Prepross}

 Speed and time difference information were widely used in sketch recognition
systems \cite{earlyprocess}. Figure \ref{fig:orignalStroke} shows an example of
an input stroke drawn by users.  Figure \ref{fig:speed2Distance} shows the graph
of time difference and the speed for the stroke in Figure
\ref{fig:orignalStroke}. It was noted that time difference provides more
distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31}
mentions that the pointing device (e.g., mouse) sampling rate is the reason for
this phenomenon. The points are sampled at regular time intervals as long as the
pointing device is moving. Therefore, there are no new detected points when the
pointing device is stationary. This leads to a nearly constant time difference
between samples while the pen is moving and large difference while the pen is
stationary (Figure \ref{fig:timediff}). Contrary to time difference information,
the speed information has a lot of noise in the data because usually  users draw
with variant speeds (Figure \ref{fig:speed2Distance}).  %Similarly, direction
%information is used as it provides better distinctive maxima for the corners
%than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures}
%shows the direction and curvature graphs for the stroke drawn in Figure
\ref{fig:orignalStroke}

 \begin{figure}[h]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system.
	\label{fig:orignalStroke}
\end{figure}

 \begin{figure}
	\centering
			\subfigure[Speed versus Displacement Graph]{
\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference  versus Displacement
Graph]{\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}  a) speed graph has lot of
noise in the data with respect to b) time difference.   The black points denote
the location of the possible dominant points $P_{pd}$. Black rectangles
represent regions $R_i$ that in a) Are lower than the threshold. b) Are above
the threshold.
	\label{fig:speed2Distance}
\end{figure}

Dominant points are characterized by low speed values, high curvature and
direction values. However, the dominant points from the curvature information
may not be the same points as the dominant points extracted from the speed
information. Hence, we compute time difference, direction, speed and curvature
of each point along a stroke to ensure that no dominant point is overlooked in
this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the
time difference between two points and $s$ is the length between them. The
direction $d$ calculated as the angle between vectors $\overrightarrow {P_{i -
1} P_i }$ and the $x-axis$ and the curvature is considered as the change in
direction $d$ with respect to length $s$ i.e. $c= \Delta d/\Delta s$.

After the system computes the speed, time difference and curvature information,
it proceeds to extracts the points that have low velocity and high curvature.
However, using simple differentiation to detect local extreme points results in
false points due to the non smooth curves. Hence, our system adopts a process
presented by \cite{earlyprocess}, where the mean of each of the four calculated
information is calculated. This mean is taken as the threshold \textit{th} which
is used to separate the curve into regions ($R_i$) (Figure\ref{fig:timediff});
each region $R_i$ is defined as a range of points, where the curve values are
either above or below the threshold \textit{th}. Those regions are further
processed to find the maximum point $Max(R_i)$ of each region $R_i$. The stroke
points $p_i(x,y)$ that correspond to those maximum values are labeled as
\textit{possible dominant points} $P_{pd}$. The system repeats this process for
curvature, time difference and speed information. All the points labeled as
possible dominant points $P_{pd}$ are saved into a single array. Figure
\ref{fig:ppd999} shows the particles labeled as Possible dominate point $P_{pd}$
in the preprocessing phase. It is clear that some of the $P_{pd}$ points are
redundant as specified in the preprocessing stage. % (as shown are redundant)
\begin{figure}
	\centering
\includegraphics[scale=0.7]{images/ppd.jpg}
	\caption{Possible Dominante Points} An illustration of stroke points and
the expected \textit{possible dominant points} $P_{pd}$.
	%\label{fig:ppd}
	\label{fig:ppd999}
\end{figure}

 %Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon.
% % In the presented system, we computed time difference, direction, speed and
% curvature of each point along a stroke. The speed is calculated as $v=\Delta
% s/\Delta t$ where $t$ is the time difference between two points and $s$ is the
% length between them. The direction $d$ is calculated as the angle between two
% vectors and curvature is considered as the change in direction with respect to
% length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower
% speed values and higher curvature and direction values.

% After the system computes all the speed, time difference and curvature
% information it proceeds to detect the points with low velocity and high
% curvatures. Using simple differentiation to detect local extreme points resulted
% in false points due to the non smooth curves. Hence, the system adopted a
% process presented by \cite{earlyprocess}, where the mean of the curve is
% calculated. Then a threshold \textit{th} is used to separate the curve into
% regions $Region_i$; each region $Region_i$ is defined as a range of points,
% where the curve values are either above or below the threshold \textit{th}.
% Those regions are further processed to find the maximum point $Max(Region_i)$ of
% each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those
% maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The
% system repeats this process for curvature, time difference and speed
% information. All the points labeled as possible dominant points $P_{pd}$ are
% saved into a single array. Figure \ref{fig:LabelsPPD} shows the particles
% labeled as Possible dominant point $P_{pd}$ by the preprocessing, it is noticed
% the redundancy of some $P_{pd}$ points. % (as shown are redundant)

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives.
As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke
points into a curve or an ellipse using a minimum square error fitting algorithm
\cite{ellipsefit}. The ellipse fitting step helps to prevent the system from
over segmenting the stroke into multiple lines or curves if the input stroke is
an ellipse. If the stroke proved to be an ellipse arc then the segmentation
process ends and the system proceeds to the next step. Otherwise, the stroke is
passed to two further segmentation algorithms that divide the stroke to either
lines or lines and curves. After two different segmentations are generated, the
system chooses the segmentation with the minimum error. The following sections
describe in details the ellipse detection algorithm and the two segmentation
algorithms used to divide the stroke.

 \begin{figure}
	\centering
		\includegraphics[scale=0.48]{blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting}

@@

	 \includegraphics[scale=0.7]{pso1.jpg}
	\caption{DPSO algorithm encoding}
	\label{fig:pso1}
\end{figure}

Two DPSO algorithms are used to generate different segmentations for each stroke
the user draws. For each segmentation generated, an error is evaluated. The
segmentation with the minimum error value is chosen as the best stroke
segmentation (Figure \ref{fig:segblock}). The problem definition is the same in
both algorithms but they differ in the method they use to compute fitness and
error functions.
\begin{description}
	\item[ Problem definition:] The input stroke $S$  with $N$ points can be
represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is
the location of the point $i$. The swarm algorithms consist of $M$ agents which
are represented by the set $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.}
\right\}$ where $P_i$ is a single solution particle from the solution space.
Each particle decodes the problem using a binary array with the same length $N$
as the input stroke.

@@

	\item[Fitness function:] The fitness function and error calculation are
different in each of the two \textit{DPSO} algorithms.
	\end{description}
\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
%\\
The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The
graphical meaning of the error is shown in Figure\ref{fig:pso1}.%\ref{fig:LabelsPSO}.
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive
set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$.
The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to
point $x_j$, $M$ is the number of dominant points in this solution as generated
by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is
computed as the sum of squared perpendicular distance from every point along the
arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed
using equation %\ref{eq:fitnessSwarm1}
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation} where $N$ is the number of points in the stroke, $D$ is the
number of points in the solution that was previously labeled as a possible
dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the
error threshold. It should be noticed that when the error is larger than the
threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness
value of the solution. Otherwise the system favors the lower number of vertices.

@@

\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes
is grouped together along with their segmentation as input to the feature
extraction process. A composite set of features is used to generate a single
feature vector. The following list describes each feature in details.
%\begin{description}
\subsubsection{Structural and geometrical Features (FS1)}
  They are features that define the structure of the geometrical symbol such as;
 \begin{itemize}
	 \item \emph{Segments:} The number of segments in the symbol.
	 \item \emph{Strokes:} The number of strokes or partial strokes that
created the symbols.
		\item  \emph{Primitives:} The number of primitives in the
symbol. This feature helps when identifying  symbols with mixed geometric
primitives such as cylinders and callouts.
		\item \emph{Curves:} The number of curves or ellipses in the
symbol.
		\item \emph{Lines:} The number of lines in the symbol.
		\item \emph{Perpendicular}\emph{lines:} The number of
perpendicular lines.
		\item \emph{Parallel}\emph{lines:} The number of parallel lines.
		\item \emph{Intersections:} The number of intersection between
lines and curves.
		\item \emph{T intersections:} The number of T intersections.
		\item \emph{L intersections:} The number of L intersections.
		\item \emph{X intersections:} The number of X intersections.
